---
title: "Course Project: Shiny Application and Reproducible Pitch"
author: "Celine Barlier"
date: "8/28/2021"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is the assignment for the course project (week 4) of the course Developing Data Product of Coursera. It consisted in the development of a Shiny application and this reproducible pitch.

I developed a Shiny Application using a large scale weather event in USA (2016-2020) dataset to plot interactive demographic maps.

**Please note that the shiny app takes initially about 1 min to load and show the map.**

Link to the Shiny Application: https://cbarlier.shinyapps.io/ShinyAppWeatherUSA/
Link to the code: https://github.com/BarlierC/DevDataProduct

## Shiny App

The shiny app display weather event frequencies depending on the type of event, severity and year(s) using the data from: https://smoosavi.org/datasets/lstw

Specific columns have been selected (shown below), a column Year has been created by extracting the specific information from the StartTime.UTC. column and a column oc (occurence) used to compute the frequency.

```{r}
df = readRDS("C:/Users/celin/Desktop/CourseraProject/ShinyApp/WeatherEvents_Jan2016-Dec2020.rds")
head(df)
```

## UI.R code
```{r eval=FALSE}
library(plotly)

ui <- fluidPage(
  
  # App title ----
  titlePanel("Weather Events in USA"),
  
  # Sidebar layout with input and output definitions ----
  sidebarLayout(
    
    # Sidebar panel for inputs ----
    sidebarPanel(
      helpText("Create demographic maps of weather events that happened in the USA from 2016 to 2020."),

        selectInput("var", 
                  label = "Weather event",
                  choices = c("Snow", 
                              "Fog",
                              "Cold", 
                              "Storm",
                              "Rain",
                              "Precipitation",
                              "Hail"),
                  selected = "Snow"
        ),
      
        checkboxGroupInput("checkGroup", 
                         h3("Severity"), 
                         choices = list("Light" = "Light", 
                                        "Moderate" = "Moderate", 
                                        "Severe" = "Severe",
                                        "Heavy" = "Heavy",
                                        "Unknown" = "Unknown"),
                         selected = c("Light","Moderate","Severe","Heavy","Unknown")
        ),
        
        sliderInput("slider", h3("Year(s)"),
                      min = 2016, max = 2020, value = c(2016,2020), step = 1
        )
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      
      # Output: Map displaying event frequency ----
      plotlyOutput("mapFreq")
      
    )
  )
)
```


## server.R code
```{r eval=FALSE}
library(plotly)
library(stringr)

#Load weather data
df = readRDS("WeatherEvents_Jan2016-Dec2020.rds")

#Add occurrence for later on
df$oc <- rep(1,nrow(df))

# Define server logic ----
server <- function(input, output) {
  
  # USA map displaying selected event frequency ----
  output$mapFreq <- renderPlotly({
    
    #Select event based on input: event type
    dfEvent <- df[which(df$Type == input$var),]
    #Select event based on input: severity/ies
    dfEvent <- dfEvent[which(dfEvent$Severity %in% c(input$checkGroup)),]
    #Select event based on input: year(s)
    dfEvent <- dfEvent[which(dfEvent$Year %in% c(input$slider)),]
    
    #If data to show
    if(nrow(dfEvent) > 0){
      #Compute frequency = sum of events
      dfFreq <- aggregate(dfEvent$oc,by=list("State"=dfEvent$State),sum)
      colnames(dfFreq) <- c("State","Freq")
      Frequency <- dfFreq$Freq
      
      #Element for map
      plot_geo() %>%
        add_trace(
          z = ~Frequency, text = dfFreq$State,
          locations = dfFreq$State, 
          locationmode = 'USA-states'
        ) %>%
        layout(geo = list(
          scope = 'usa',
          projection = list(type = 'albers usa'),
          lakecolor = toRGB('white')
        ))
    }else{
      plotly_empty() %>%
      layout(title = "No data to show with selected parameters") 
    }

  })
  
}
```
